---
title: "Guiana 2023"
author: "Megan Reich"
date: "08/02/2022"
output: 
  html_document:
    theme: paper
    toc: true
    toc_float: true
---
```{r setup}
knitr::opts_chunk$set(echo = TRUE)
```
# Purpose
Code for the isotope-based geographic assignment section of:  
Suchan, T., Bataille, C.P., Reich, M.S., Toro-Delgado, E. Vila, R., Pierce, N.E., Talavera, G. (2024). A trans-oceanic flight of over 4,200 km by painted lady butterflies. Nature Communications.  
  
  
Isotope-based geographic assignment using hydrogen and strontium isotopes of 3 painted ladies captured in French Guiana by Gerard Talavera in October 2013. The hydrogen isotope values were measured in Utah by IsoForensics in June 2017. The strontium isotope ratios were measured at UBC by Megan Reich in December 2021.

# Setup
```{r message=FALSE, warning=FALSE}
# packages
if(!require(osfr)) install.packages("osfr"); library(osfr)
if (!require(tidyverse)) install.packages(tidyverse); library(tidyverse)
if (!require(terra)) install.packages(terra); library(terra)
if (!require(assignR)) install.packages(assignR); library(assignR)
if (!require(rnaturalearth)) install.packages(rnaturalearth); library(rnaturalearth)
if (!require(viridis)) install.packages(viridis); library(viridis)
if (!require(lattice)) install.packages(lattice); library(lattice)
if (!require(unikn)) install.packages(unikn); library(unikn)
if (!require(ggpubr)) install.packages(ggpubr); library(ggpubr)
if (!require(tidyterra)) install.packages(tidyterra); library(tidyterra)
if (!require(sf)) install.packages(sf); library(sf)
```
```{r}
#make folders in home directory
dir.create("output")
dir.create("output/H")
dir.create("output/H_Sr")
dir.create("output/Sr")
dir.create("isoscapes")
```
# samples
```{r}
###Guiana isotope data
iso.data <- read.csv("unknown/IsoData_Guiana.csv") # tissue isotope data
```
```{r message=FALSE, warning=FALSE}
# projections
WGS84 <- "+init=epsg:4326"
eckert<-"+proj=eck4 +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 +units=m +no_defs"

## define study extent 
box<-c(xmin=-10470815, ymin=-1314163, xmax= 4682185,  ymax=8259837) 
### Create a country shapefile
extent1<-ne_countries(returnclass = "sf")
extent1<-sf::st_transform(extent1,crs=eckert)
extent1 <- sf::st_buffer(extent1, dist=0)
extent3 <- sf::st_crop(x=extent1, y=box)
sf::st_write(obj = extent3, dsn = "isoscapes/extent3.shp", delete_dsn = TRUE)
```
The unknown origin data was measured in Utah (IsoForensics Inc) and is on Magozzi et al (2021) reference code: UT_H_2
```{r}
#add ref scale
iso.data$d2H_cal<-"UT_H_2"
#dataframe as required
samples<-iso.data %>%
  dplyr::select(c(d2H = d2h,d2H.sd = d2h_err,d2H_cal))
#transform to VSMOW
sample_trans<-assignR::refTrans(samples=samples, marker="d2H", ref_scale = "VSMOW_H")
```
```{r}
print(cbind(iso.data[,1:3], sample_trans$data))
```
# Isoscapes
## Strontium isoscape - part 1
```{r}
# Download Bataille et al 2020 Global bioavailable 87Sr/86Sr isoscapes from the Open Science Foundation
#OSF repo
sr_rasters <- osf_retrieve_node("pwdvn")
osf_ls_files(sr_rasters)

sr_rasters %>%
  osf_ls_files(path ="Stontium isoscape/", pattern = ".tif")%>%
  osf_download(path="./isoscapes/")
```

```{r eval=FALSE, warning=FALSE}
###load isoscapes Sr and uncertainty from Bataille et al 2020
rf_sr<-rast("isoscapes/rf_plantsoilmammal1.tif")
rf_sr_err <-rast("isoscapes/srse.tif")

###Crop Sr isoscape to study area
Sr_mean <- terra::crop(rf_sr, extent3)
Sr_sd <- terra::crop(rf_sr_err, extent3)
```

continued below

## Hydrogen isoscape
### precipitation isoscape
```{r eval=FALSE, warning=FALSE}
###get Global mean-annual precipitation H and O isotope values
d2h_world<-assignR::getIsoscapes(isoType = "GlobalPrecipMA", timeout = 1200)
d2h_world<-subset(d2h_world,1:2)
```
Format: GeoTIFF
Spatial extent: Global land surface
Spatial resolution: 5' x 5'
Temporal extent: Climatological, monthly or annual average (preciptation amount-weighted)
Projection: WGS84 unprojected
Layers:
*d2h_MA: Mean predicted annual-average d2H values
*d2h_se_MA: Standard error of the mean predicted annual-average d2H values
*d2h_01...: Mean predicted monthly d2H values for January (01) through December (12)
*d2h_se_01...: Standard error of the mean predicted monthly d2H values for January (01) through December (12)
*d18o_MA: Mean predicted annual-average d18O values
*d18o_se_MA: Standard error of the mean predicted annual-average d18O values
*d18o_01...: Mean predicted monthly d18O values for January (01) through December (12)
*d18o_se_01...: Standard error of the mean predicted monthly d18O values for January (01) through December (12)
Source: Bowen, G. J., & Revenaugh, J. (2003). Interpolating the isotopic composition of modern meteoric precipitation. Water resources research, 39(10). 
Source: Bowen, G. J., Wassenaar, L. I., & Hobson, K. A. (2005). Global application of stable hydrogen and oxygen isotopes to wildlife forensics. Oecologia, 143(3), 337-348.

### calibration data
#### Hobson et al 2019
```{r}
# Get calibration data
#Hobson 2019
caldata<-assignR::subOrigData(marker="d2H", taxon="Danaus plexippus", ref_scale="VSMOW_H") #ref_scale:	character string. Text identifier for reference scale to which all isotope values will be transformed. 
caldata
#Result: 150 samples from 31 sites Hobseon et al 2019
```
#### Ghouri et al 2023
```{r}
caldata_Ghouri<-read.csv("isoscapes/knownorigin_butterfly_2023_final_v8.csv",fileEncoding="UTF-8-BOM")
caldata_Ghouri2<-subset(caldata_Ghouri, select=c("d2H.average","d2H.sd","Lon", "Lat"))

# no meaningful difference (~0.02 per mil) with reftrans correction so used original values
caltrans<-caldata_Ghouri2 %>%
  dplyr::select(c(d2H = d2H.average,d2H.sd = d2H.sd))
caltrans$d2H_cal<-"EC_H_9"
#transform to VSMOW
caltrans2<-refTrans(samples=caltrans, marker="d2H", ref_scale = "VSMOW_H")
print(cbind(caltrans,caltrans2$data))
```
```{r}
# if the site fall within an empty cell (usually near the coast), function to move the coordinates of the sampling site to the nearest non-NA cell 
nearest_non_na <- function(coordinates, raster) {
  # Convert the coordinates data.frame to an sf points object
  points <- st_as_sf(coordinates, coords = c("Lon", "Lat"), crs = st_crs(raster))
  
  # Extract raster values under the points
  values <- terra::extract(raster, points,method="simple")
  
  for(i in which(is.na(values[2]))){    #if a value is NA, move it
      point <- points[i,]
      # Distance does not work when the crs is "". needs an exception
      distRaster <- terra::distance(raster, point)
      distRaster <- terra::mask(distRaster, raster)
      vmin <- terra::where.min(distRaster)
      coordinates[i,3:4] = terra::xyFromCell(distRaster, vmin[2])
    }
  
  return(coordinates)
}
```
```{r}
caldata_Ghouri3 <- nearest_non_na(caldata_Ghouri2, d2h_world[[1]])
```
#### Combine
```{r}
Hobson<-as.data.frame(caldata$data)
Hobson_xy <- as.data.frame(geom(caldata$data))
Hobson <- cbind(Hobson, Hobson_xy)
Hobson<-Hobson %>%
  dplyr::select(d2H, d2H.sd, x, y)
Hobson$set<-"Hobson"
```
```{r}
Ghouri<-caldata_Ghouri3 %>%
  dplyr::select(d2H.average,d2H.sd,Lon,Lat)%>%
  rename(d2H =d2H.average, x = Lon, y =Lat)
Ghouri$set<-"Ghouri"
```
```{r}
Cal <- rbind(Hobson, Ghouri)

#terra
Cal_HobGho<-vect(Cal, geom =c("x", "y"))
crs(Cal_HobGho) <- WGS84
```

### calibrate H isoscape

```{r}
#calibrate isoscape
d2Hw_HobGho<-calRaster(known=Cal_HobGho, isoscape=d2h_world, interpMethod = 2, outDir="isoscapes")

###Reproject H isoscape to Sr isoscape
d2Hw_HobGho_p<-terra::project(x = d2Hw_HobGho$isoscape.rescale, y= Sr_mean,method="bilinear", align=T)

###Cut wing H isoscape to same extent as Sr isoscape
d2Hw_HobGho_p<-crop(d2Hw_HobGho_p,Sr_mean)
d2Hw_HobGho_p<-mask(d2Hw_HobGho_p, Sr_mean)

#save
writeRaster(d2Hw_HobGho_p[[1]], "isoscapes/d2Hw_mean_HobGho.tif", overwrite=T)
writeRaster(d2Hw_HobGho_p[[2]], "isoscapes/d2Hw_sd_HobGho.tif", overwrite=T)
```
```{r}
#bring in final H isoscape
d2Hw_mean_HobGho<-rast("isoscapes/d2Hw_mean_HobGho.tif")
d2Hw_sd_HobGho<-rast("isoscapes/d2Hw_sd_HobGho.tif")
d2Hw_HobGho_p<-c(d2Hw_mean_HobGho,d2Hw_sd_HobGho)
```

## Strontium isoscape - part 2
```{r}
###Cut wing Sr isoscape to same extent as H isoscape
Sr_mean<-crop(Sr_mean,d2Hw_HobGho_p[[1]]) #hydrogen
Sr_mean<-terra::mask(Sr_mean, d2Hw_HobGho_p[[1]])
Sr_sd<-crop(Sr_sd,d2Hw_HobGho_p[[1]]) #hydrogen
Sr_sd2<-mask(Sr_sd, d2Hw_HobGho_p[[1]])

#save
writeRaster(Sr_mean, "isoscapes/Sr_Bataille2020_mean_V2.tif", overwrite=T)
writeRaster(Sr_sd, "isoscapes/Sr_Bataille2020_sd_V2.tif", overwrite=T)
```
```{r}
#bring in final Sr isoscape
Sr_mean <- rast("isoscapes/Sr_Bataille2020_mean_V2.tif")
Sr_sd <- rast("isoscapes/Sr_Bataille2020_sd_V2.tif")
###Create stack for assignR
Sr_stack<-c(Sr_mean,Sr_sd)
```

# d2H assignment
```{r}
H_postps<-assignR::pdRaster(d2Hw_HobGho_p, unknown = sample_trans, outDir = "output/H/")

#joint
joint_H<-assignR::jointP(H_postps)
writeRaster(joint_H, "output/H/Joint_H.tif")
```
```{r}
H_973<-rast("output/H/1_like.tif")
H_974<-rast("output/H/2_like.tif")
H_975<-rast("output/H/3_like.tif")
joint_H<-rast("output/H/Joint_H.tif")
```
## graticules and locations
```{r}
grat <- sf::st_graticule(lon=seq(-160,90, 40),
                      lat = seq(0,80, 20),
                      ndiscr = 5000) %>%
  vect() %>%
  project(eckert) 
```
```{r}
# capture location
indxy <- sf::st_as_sf(iso.data, coords = c("long", "lat"), crs=WGS84) #WGS84
indxy <- sf::st_transform(indxy, crs = eckert)
```
## plot
```{r}
H_973 <- terra::setMinMax(H_973, force =T) #omg it doesnt auto display the actual min and max
Uniques<-terra::minmax(H_973)
Uniques.max <-Uniques["max","lyr.1"]
Uniques.min <-Uniques["min","lyr.1"]
#try a ggplot
H_973_plot <- ggplot()+
   geom_spatraster(data = H_973) +
   scale_fill_gradient2(limits = c(Uniques.min, Uniques.max), low = "#F2F2F2",mid= "#21908CFF", high = "#FDE725FF", midpoint=mean(c(Uniques.max,Uniques.min)),
    na.value = NA, breaks = seq(Uniques.min, Uniques.max, length.out = 5), labels = scales::label_scientific(digits =2)) +
  labs(fill= "Probability")+
   geom_sf(data=extent3,color="grey30",fill="transparent", lwd=0.3)+ #countries
   geom_sf(data=indxy, color="red", pch=1, cex=.9)+
   geom_spatvector(data=grat, color=alpha("grey60", 0.5)) +
   coord_sf(xlim = c(-10470815, 4682185),ylim = c(-1314163, 8259837), expand=F)+
   scale_x_continuous(breaks = seq(-160, 70, by = 40)) +
   scale_y_continuous(breaks = seq(0, 80, by = 20))+
   theme_minimal()+
   theme(plot.margin = margin(0, 0, 0, 0, "cm"),
         legend.margin=margin(0,0,.2,0),
         plot.background = element_rect(fill=NA, color=NA),
         panel.border=element_rect(fill=NA, color="grey30"),
         legend.background = element_rect(fill = NA, color=NA),
         legend.key.height = unit(1, "cm"),
         axis.title.x = element_blank(), 
     axis.title.y = element_blank())
ggsave(plot=H_973_plot, filename="output/H/H_973_plot.png", device="png", height=4, width=8, units="in",bg="white")
```
```{r}
H_974 <- terra::setMinMax(H_974, force =T) #omg it doesnt auto display the actual min and max
Uniques<-terra::minmax(H_974)
Uniques.max <-Uniques["max","lyr.1"]
Uniques.min <-Uniques["min","lyr.1"]
#try a ggplot
H_974_plot <- ggplot()+
   geom_spatraster(data = H_974) +
   scale_fill_gradient2(limits = c(Uniques.min, Uniques.max), low = "#F2F2F2",mid= "#21908CFF", high = "#FDE725FF", midpoint=mean(c(Uniques.max,Uniques.min)),
    na.value = NA, breaks = seq(Uniques.min, Uniques.max, length.out = 5), labels = scales::label_scientific(digits =2)) +
   labs(fill= "Probability")+
   geom_sf(data=extent3,color="grey30",fill="transparent", lwd=0.3)+ #countries
   geom_sf(data=indxy, color="red", pch=1, cex=.9)+
   geom_spatvector(data=grat, color=alpha("grey60", 0.5)) +
   coord_sf(xlim = c(-10470815, 4682185),ylim = c(-1314163, 8259837), expand=F)+
   scale_x_continuous(breaks = seq(-160, 70, by = 40)) +
   scale_y_continuous(breaks = seq(0, 80, by = 20))+
   theme_minimal()+
   theme(plot.margin = margin(0, 0, 0, 0, "cm"),
         legend.margin=margin(0,0,.2,0),
         plot.background = element_rect(fill=NA, color=NA),
         panel.border=element_rect(fill=NA, color="grey30"),
         legend.background = element_rect(fill = NA, color=NA),
         legend.key.height = unit(1, "cm"),
         axis.title.x = element_blank(), 
     axis.title.y = element_blank())
ggsave(plot=H_974_plot, filename="output/H/H_974_plot.png", device="png", height=4, width=8, units="in",bg="white")
```
```{r}
H_975 <- terra::setMinMax(H_975, force =T) #omg it doesnt auto display the actual min and max
Uniques<-terra::minmax(H_975)
Uniques.max <-Uniques["max","lyr.1"]
Uniques.min <-Uniques["min","lyr.1"]
#try a ggplot
H_975_plot <- ggplot()+
   geom_spatraster(data = H_975) +
   scale_fill_gradient2(limits = c(Uniques.min, Uniques.max), low = "#F2F2F2",mid= "#21908CFF", high = "#FDE725FF", midpoint=mean(c(Uniques.max,Uniques.min)),
    na.value = NA, breaks = seq(Uniques.min, Uniques.max, length.out = 5), labels = scales::label_scientific(digits =2)) +
   labs(fill= "Probability")+
   geom_sf(data=extent3,color="grey30",fill="transparent", lwd=0.3)+ #countries
   geom_sf(data=indxy, color="red", pch=1, cex=.9)+
   geom_spatvector(data=grat, color=alpha("grey60", 0.5)) +
   coord_sf(xlim = c(-10470815, 4682185),ylim = c(-1314163, 8259837), expand=F)+
   scale_x_continuous(breaks = seq(-160, 70, by = 40)) +
   scale_y_continuous(breaks = seq(0, 80, by = 20))+
   theme_minimal()+
   theme(plot.margin = margin(0, 0, 0, 0, "cm"),
         legend.margin=margin(0,0,.2,0),
         plot.background = element_rect(fill=NA, color=NA),
         panel.border=element_rect(fill=NA, color="grey30"),
         legend.background = element_rect(fill = NA, color=NA),
         legend.key.height = unit(1, "cm"),
         axis.title.x = element_blank(), 
     axis.title.y = element_blank())
ggsave(plot=H_975_plot, filename="output/H/H_975_plot.png", device="png", height=4, width=8, units="in",bg="white")
```
```{r}
joint_H <- terra::setMinMax(joint_H, force =T) #omg it doesnt auto display the actual min and max
Uniques<-terra::minmax(joint_H)
Uniques.max <-Uniques["max","Joint_Probability"]
Uniques.min <-Uniques["min","Joint_Probability"]
#try a ggplot
joint_H_plot <- ggplot()+
   geom_spatraster(data = joint_H) +
   scale_fill_gradient2(limits = c(Uniques.min, Uniques.max), low = "#F2F2F2",mid= "#21908CFF", high = "#FDE725FF", midpoint=mean(c(Uniques.max,Uniques.min)),
    na.value = NA, breaks = seq(Uniques.min, Uniques.max, length.out = 5), labels = scales::label_scientific(digits =2)) +
   labs(fill= "Joint probability")+
   geom_sf(data=extent3,color="grey30",fill="transparent", lwd=0.3)+ #countries
   geom_sf(data=indxy, color="red", pch=1, cex=.9)+
   geom_spatvector(data=grat, color=alpha("grey60", 0.5)) +
   coord_sf(xlim = c(-10470815, 4682185),ylim = c(-1314163, 8259837), expand=F)+
   scale_x_continuous(breaks = seq(-160, 70, by = 40)) +
   scale_y_continuous(breaks = seq(0, 80, by = 20))+
   theme_minimal()+
   theme(plot.margin = margin(0, 0, 0, 0, "cm"),
         legend.margin=margin(0,0,.2,0),
         plot.background = element_rect(fill=NA, color=NA),
         panel.border=element_rect(fill=NA, color="grey30"),
         legend.background = element_rect(fill = NA, color=NA),
         legend.key.height = unit(1, "cm"),
         axis.title.x = element_blank(), 
     axis.title.y = element_blank())
ggsave(plot=joint_H_plot, filename="output/H/joint_H_plot.png", device="png", height=4, width=8, units="in",bg="white")
```

# 87Sr/86Sr assignment
```{r eval=FALSE}
###dataset prepared for assignR
iso.data_Sr<-subset(iso.data, select=c("ID","Sr","Sr_err"))
```
```{r eval=FALSE}
###pd Sr
Sr_pd<-assignR::pdRaster(Sr_stack, unknown = iso.data_Sr, outDir = "output/Sr/")
```
```{r eval=FALSE}
###Joint Sr prob
Joint_Sr<-assignR::jointP(Sr_pd)
writeRaster(Joint_Sr, "output/Sr/Joint_Sr.tif")
```
```{r eval=FALSE, warning=FALSE}
Sr_973<-rast("output/Sr/RVcoll14M973_like.tif")
Sr_974<-rast("output/Sr/RVcoll14M974_like.tif")
Sr_975<-rast("output/Sr/RVcoll14M975_like.tif")
Joint_Sr<-rast("output/Sr/Joint_Sr.tif")
```
```{r}
Sr_973 <- terra::setMinMax(Sr_973, force =T) #omg it doesnt auto display the actual min and max
Uniques<-terra::minmax(Sr_973)
Uniques.max <-Uniques["max","rf_plantsoilmammal1"]
Uniques.min <-Uniques["min","rf_plantsoilmammal1"]
#try a ggplot
Sr_973_plot <- ggplot()+
   geom_spatraster(data = Sr_973) +
   scale_fill_gradient2(limits = c(Uniques.min, Uniques.max), low = "#F2F2F2",mid= "#21908CFF", high = "#FDE725FF", midpoint=mean(c(Uniques.max,Uniques.min)),
    na.value = NA, breaks = seq(Uniques.min, Uniques.max, length.out = 5), labels = scales::label_scientific(digits =2)) +
   labs(fill= "Probability")+
   geom_sf(data=extent3,color="grey30",fill="transparent", lwd=0.3)+ #countries
   geom_sf(data=indxy, color="red", pch=1, cex=.9)+
   geom_spatvector(data=grat, color=alpha("grey60", 0.5)) +
   coord_sf(xlim = c(-10470815, 4682185),ylim = c(-1314163, 8259837), expand=F)+
   scale_x_continuous(breaks = seq(-160, 70, by = 40)) +
   scale_y_continuous(breaks = seq(0, 80, by = 20))+
   theme_minimal()+
   theme(plot.margin = margin(0, 0, 0, 0, "cm"),
         legend.margin=margin(0,0,.2,0),
         plot.background = element_rect(fill=NA, color=NA),
         panel.border=element_rect(fill=NA, color="grey30"),
         legend.background = element_rect(fill = NA, color=NA),
         legend.key.height = unit(1, "cm"),
         axis.title.x = element_blank(), 
     axis.title.y = element_blank())
ggsave(plot=Sr_973_plot, filename="output/Sr/Sr_973_plot.png", device="png", height=4, width=8, units="in",bg="white")
```
```{r}
Sr_974 <- terra::setMinMax(Sr_974, force =T) #omg it doesnt auto display the actual min and max
Uniques<-terra::minmax(Sr_974)
Uniques.max <-Uniques["max","rf_plantsoilmammal1"]
Uniques.min <-Uniques["min","rf_plantsoilmammal1"]
#try a ggplot
Sr_974_plot <- ggplot()+
   geom_spatraster(data = Sr_974) +
   scale_fill_gradient2(limits = c(Uniques.min, Uniques.max), low = "#F2F2F2",mid= "#21908CFF", high = "#FDE725FF", midpoint=mean(c(Uniques.max,Uniques.min)),
    na.value = NA, breaks = seq(Uniques.min, Uniques.max, length.out = 5), labels = scales::label_scientific(digits =2)) +
   labs(fill= "Probability")+
   geom_sf(data=extent3,color="grey30",fill="transparent", lwd=0.3)+ #countries
   geom_sf(data=indxy, color="red", pch=1, cex=.9)+
   geom_spatvector(data=grat, color=alpha("grey60", 0.5)) +
   coord_sf(xlim = c(-10470815, 4682185),ylim = c(-1314163, 8259837), expand=F)+
   scale_x_continuous(breaks = seq(-160, 70, by = 40)) +
   scale_y_continuous(breaks = seq(0, 80, by = 20))+
   theme_minimal()+
   theme(plot.margin = margin(0, 0, 0, 0, "cm"),
         legend.margin=margin(0,0,.2,0),
         plot.background = element_rect(fill=NA, color=NA),
         panel.border=element_rect(fill=NA, color="grey30"),
         legend.background = element_rect(fill = NA, color=NA),
         legend.key.height = unit(1, "cm"),
         axis.title.x = element_blank(), 
     axis.title.y = element_blank())
ggsave(plot=Sr_974_plot, filename="output/Sr/Sr_974_plot.png", device="png", height=4, width=8, units="in",bg="white")
```
```{r}
Sr_975 <- terra::setMinMax(Sr_975, force =T) #omg it doesnt auto display the actual min and max
Uniques<-terra::minmax(Sr_975)
Uniques.max <-Uniques["max","rf_plantsoilmammal1"]
Uniques.min <-Uniques["min","rf_plantsoilmammal1"]
#try a ggplot
Sr_975_plot <- ggplot()+
   geom_spatraster(data = Sr_975) +
   scale_fill_gradient2(limits = c(Uniques.min, Uniques.max), low = "#F2F2F2",mid= "#21908CFF", high = "#FDE725FF", midpoint=mean(c(Uniques.max,Uniques.min)),
    na.value = NA, breaks = seq(Uniques.min, Uniques.max, length.out = 5), labels = scales::label_scientific(digits =2)) +
   labs(fill= "Probability")+
   geom_sf(data=extent3,color="grey30",fill="transparent", lwd=0.3)+ #countries
   geom_sf(data=indxy, color="red", pch=1, cex=.9)+
   geom_spatvector(data=grat, color=alpha("grey60", 0.5)) +
   coord_sf(xlim = c(-10470815, 4682185),ylim = c(-1314163, 8259837), expand=F)+
   scale_x_continuous(breaks = seq(-160, 70, by = 40)) +
   scale_y_continuous(breaks = seq(0, 80, by = 20))+
   theme_minimal()+
   theme(plot.margin = margin(0, 0, 0, 0, "cm"),
         legend.margin=margin(0,0,.2,0),
         plot.background = element_rect(fill=NA, color=NA),
         panel.border=element_rect(fill=NA, color="grey30"),
         legend.background = element_rect(fill = NA, color=NA),
         legend.key.height = unit(1, "cm"),
         axis.title.x = element_blank(), 
     axis.title.y = element_blank())
ggsave(plot=Sr_975_plot, filename="output/Sr/Sr_975_plot.png", device="png", height=4, width=8, units="in",bg="white")
```
```{r}
Joint_Sr <- terra::setMinMax(Joint_Sr, force =T) #omg it doesnt auto display the actual min and max
Uniques<-terra::minmax(Joint_Sr)
Uniques.max <-Uniques["max","Joint_Probability"]
Uniques.min <-Uniques["min","Joint_Probability"]
#try a ggplot
Joint_Sr_plot <- ggplot()+
   geom_spatraster(data = Joint_Sr) +
   scale_fill_gradient2(limits = c(Uniques.min, Uniques.max), low = "#F2F2F2",mid= "#21908CFF", high = "#FDE725FF", midpoint=mean(c(Uniques.max,Uniques.min)),
    na.value = NA, breaks = seq(Uniques.min, Uniques.max, length.out = 5), labels = scales::label_scientific(digits =2)) +
   labs(fill= "Joint probability")+
   geom_sf(data=extent3,color="grey30",fill="transparent", lwd=0.3)+ #countries
   geom_sf(data=indxy, color="red", pch=1, cex=.9)+
   geom_spatvector(data=grat, color=alpha("grey60", 0.5)) +
   coord_sf(xlim = c(-10470815, 4682185),ylim = c(-1314163, 8259837), expand=F)+
   scale_x_continuous(breaks = seq(-160, 70, by = 40)) +
   scale_y_continuous(breaks = seq(0, 80, by = 20))+
   theme_minimal()+
   theme(plot.margin = margin(0, 0, 0, 0, "cm"),
         legend.margin=margin(0,0,.2,0),
         plot.background = element_rect(fill=NA, color=NA),
         panel.border=element_rect(fill=NA, color="grey30"),
         legend.background = element_rect(fill = NA, color=NA),
         legend.key.height = unit(1, "cm"),
         axis.title.x = element_blank(), 
     axis.title.y = element_blank())
ggsave(plot=Joint_Sr_plot, filename="output/Sr/Joint_Sr_plot.png", device="png", height=4, width=8, units="in",bg="white")
```

# dual d2H - 87Sr/86Sr assignment
```{r}
###prepare dataset for dual assignment in assignR
iso.data<-cbind(iso.data, sample_trans$data)
iso.data_dual<-subset(iso.data, select=c("ID","d2H","Sr"))
iso.data_dual
```
```{r}
#put the data together
dual_hobgho_stack<-isoStack(d2Hw_HobGho_p,Sr_stack)
```
```{r}
dual_pd <- assignR::pdRaster(dual_hobgho_stack, unknown = iso.data_dual, genplot = F, outDir = "output/H_Sr/")
```
```{r}
joint_d<-assignR::jointP(dual_pd)
writeRaster(joint_d, "output/H_Sr/joint_d.tif")
```
```{r}
d_973<-rast("output/H_Sr/RVcoll14M973_like.tif")
d_974<-rast("output/H_Sr/RVcoll14M974_like.tif")
d_975<-rast("output/H_Sr/RVcoll14M975_like.tif")
joint_d<-rast("output/H_Sr/joint_d.tif")
```
```{r}
d_973 <- terra::setMinMax(d_973, force =T) #omg it doesnt auto display the actual min and max
Uniques<-terra::minmax(d_973)
Uniques.max <-Uniques["max","rf_plantsoilmammal1"]
Uniques.min <-Uniques["min","rf_plantsoilmammal1"]
#try a ggplot
d_973_plot <- ggplot()+
   geom_spatraster(data = d_973) +
   scale_fill_gradient2(limits = c(Uniques.min, Uniques.max), low = "#F2F2F2",mid= "#21908CFF", high = "#FDE725FF", midpoint=mean(c(Uniques.max,Uniques.min)),
    na.value = NA, breaks = seq(Uniques.min, Uniques.max, length.out = 5), labels = scales::label_scientific(digits =2)) +
   labs(fill= "Probability")+
   geom_sf(data=extent3,color="grey30",fill="transparent", lwd=0.3)+ #countries
   geom_sf(data=indxy, color="red", pch=1, cex=.9)+
   geom_spatvector(data=grat, color=alpha("grey60", 0.5)) +
   coord_sf(xlim = c(-10470815, 4682185),ylim = c(-1314163, 8259837), expand=F)+
   scale_x_continuous(breaks = seq(-160, 70, by = 40)) +
   scale_y_continuous(breaks = seq(0, 80, by = 20))+
   theme_minimal()+
   theme(plot.margin = margin(0, 0, 0, 0, "cm"),
         legend.margin=margin(0,0,.2,0),
         plot.background = element_rect(fill=NA, color=NA),
         panel.border=element_rect(fill=NA, color="grey30"),
         legend.background = element_rect(fill = NA, color=NA),
         legend.key.height = unit(1, "cm"),
         axis.title.x = element_blank(), 
     axis.title.y = element_blank())
ggsave(plot=d_973_plot, filename="output/H_Sr/d_973_plot.png", device="png", height=4, width=8, units="in",bg="white")
```
```{r}
d_974 <- terra::setMinMax(d_974, force =T) #omg it doesnt auto display the actual min and max
Uniques<-terra::minmax(d_974)
Uniques.max <-Uniques["max","rf_plantsoilmammal1"]
Uniques.min <-Uniques["min","rf_plantsoilmammal1"]
#try a ggplot
d_974_plot <- ggplot()+
   geom_spatraster(data = d_974) +
   scale_fill_gradient2(limits = c(Uniques.min, Uniques.max), low = "#F2F2F2",mid= "#21908CFF", high = "#FDE725FF", midpoint=mean(c(Uniques.max,Uniques.min)),
    na.value = NA, breaks = seq(Uniques.min, Uniques.max, length.out = 5), labels = scales::label_scientific(digits =2)) +
   labs(fill= "Probability")+
   geom_sf(data=extent3,color="grey30",fill="transparent", lwd=0.3)+ #countries
   geom_sf(data=indxy, color="red", pch=1, cex=.9)+
   geom_spatvector(data=grat, color=alpha("grey60", 0.5)) +
   coord_sf(xlim = c(-10470815, 4682185),ylim = c(-1314163, 8259837), expand=F)+
   scale_x_continuous(breaks = seq(-160, 70, by = 40)) +
   scale_y_continuous(breaks = seq(0, 80, by = 20))+
   theme_minimal()+
   theme(plot.margin = margin(0, 0, 0, 0, "cm"),
         legend.margin=margin(0,0,.2,0),
         plot.background = element_rect(fill=NA, color=NA),
         panel.border=element_rect(fill=NA, color="grey30"),
         legend.background = element_rect(fill = NA, color=NA),
         legend.key.height = unit(1, "cm"),
         axis.title.x = element_blank(), 
     axis.title.y = element_blank())
ggsave(plot=d_974_plot, filename="output/H_Sr/d_974_plot.png", device="png", height=4, width=8, units="in",bg="white")
```
```{r}
d_975 <- terra::setMinMax(d_975, force =T) #omg it doesnt auto display the actual min and max
Uniques<-terra::minmax(d_975)
Uniques.max <-Uniques["max","rf_plantsoilmammal1"]
Uniques.min <-Uniques["min","rf_plantsoilmammal1"]
#try a ggplot
d_975_plot <- ggplot()+
   geom_spatraster(data = d_975) +
   scale_fill_gradient2(limits = c(Uniques.min, Uniques.max), low = "#F2F2F2",mid= "#21908CFF", high = "#FDE725FF", midpoint=mean(c(Uniques.max,Uniques.min)),
    na.value = NA, breaks = seq(Uniques.min, Uniques.max, length.out = 5), labels = scales::label_scientific(digits =2)) +
   labs(fill= "Probability")+
   geom_sf(data=extent3,color="grey30",fill="transparent", lwd=0.3)+ #countries
   geom_sf(data=indxy, color="red", pch=1, cex=.9)+
   geom_spatvector(data=grat, color=alpha("grey60", 0.5)) +
   coord_sf(xlim = c(-10470815, 4682185),ylim = c(-1314163, 8259837), expand=F)+
   scale_x_continuous(breaks = seq(-160, 70, by = 40)) +
   scale_y_continuous(breaks = seq(0, 80, by = 20))+
   theme_minimal()+
   theme(plot.margin = margin(0, 0, 0, 0, "cm"),
         legend.margin=margin(0,0,.2,0),
         plot.background = element_rect(fill=NA, color=NA),
         panel.border=element_rect(fill=NA, color="grey30"),
         legend.background = element_rect(fill = NA, color=NA),
         legend.key.height = unit(1, "cm"),
         axis.title.x = element_blank(), 
     axis.title.y = element_blank())
ggsave(plot=d_975_plot, filename="output/H_Sr/d_975_plot.png", device="png", height=4, width=8, units="in",bg="white")
```
```{r}
joint_d <- terra::setMinMax(joint_d, force =T) #omg it doesnt auto display the actual min and max
Uniques<-terra::minmax(joint_d)
Uniques.max <-Uniques["max","Joint_Probability"]
Uniques.min <-Uniques["min","Joint_Probability"]
#try a ggplot
joint_d_plot <- ggplot()+
   geom_spatraster(data = joint_d) +
   scale_fill_gradient2(limits = c(Uniques.min, Uniques.max), low = "#F2F2F2",mid= "#21908CFF", high = "#FDE725FF", midpoint=mean(c(Uniques.max,Uniques.min)),
    na.value = NA, breaks = seq(Uniques.min, Uniques.max, length.out = 5), labels = scales::label_scientific(digits =2)) +
   labs(fill= "Joint probability")+
   geom_sf(data=extent3,color="grey30",fill="transparent", lwd=0.3)+ #countries
   geom_sf(data=indxy, color="red", pch=1, cex=.9)+
   geom_spatvector(data=grat, color=alpha("grey60", 0.5)) +
   coord_sf(xlim = c(-10470815, 4682185),ylim = c(-1314163, 8259837), expand=F)+
   scale_x_continuous(breaks = seq(-160, 70, by = 40)) +
   scale_y_continuous(breaks = seq(0, 80, by = 20))+
   theme_minimal()+
   theme(plot.margin = margin(0, 0, 0, 0, "cm"),
         legend.margin=margin(0,0,.2,0),
         plot.background = element_rect(fill=NA, color=NA),
         panel.border=element_rect(fill=NA, color="grey30"),
         legend.background = element_rect(fill = NA, color=NA),
         legend.key.height = unit(1, "cm"),
         axis.title.x = element_blank(), 
     axis.title.y = element_blank())
ggsave(plot=joint_d_plot, filename="output/H_Sr/joint_d_plot.png", device="png", height=4, width=8, units="in",bg="white")
```
```{r}
#normalized
joint_d <- terra::setMinMax(joint_d, force =T) #omg it doesnt auto display the actual min and max
Uniques<-terra::minmax(joint_d)
Uniques.max <-Uniques["max","Joint_Probability"]
joint_d_n <- joint_d/Uniques.max
writeRaster(joint_d_n, "output/H_Sr/joint_d_plot_n.tif")
#try a ggplot
joint_d_n_plot <- ggplot()+
   geom_spatraster(data = joint_d_n) +
   scale_fill_gradient2(limits = c(0, 1), low = "#F2F2F2",mid= "#21908CFF", high = "#FDE725FF", midpoint=.5, na.value = NA, breaks = seq(0, 1, length.out = 5), labels = scales::number_format(accuracy=.1)) +
   labs(fill= "Joint probability")+
   geom_sf(data=extent3,color="grey30",fill="transparent", lwd=0.3)+ #countries
   geom_sf(data=indxy, color="red", pch=1, cex=.9)+
   geom_spatvector(data=grat, color=alpha("grey60", 0.5)) +
   coord_sf(xlim = c(-10470815, 4682185),ylim = c(-1314163, 8259837), expand=F)+
   scale_x_continuous(breaks = seq(-160, 70, by = 40)) +
   scale_y_continuous(breaks = seq(0, 80, by = 20))+
   theme_minimal()+
   theme(plot.margin = margin(0, 0, 0, 0, "cm"),
         legend.margin=margin(0,0,.2,0),
         plot.background = element_rect(fill=NA, color=NA),
         panel.border=element_rect(fill=NA, color="grey30"),
         legend.background = element_rect(fill = NA, color=NA),
         legend.key.height = unit(1, "cm"),
         axis.title.x = element_blank(), 
     axis.title.y = element_blank())
ggsave(plot=joint_d_n_plot, filename="output/H_Sr/joint_d_plot_n.png", device="png", height=4, width=8, units="in",bg="white")
```

# Additional figures
## Isoscapes
```{r}
d2Hw_mean_HobGho <- terra::setMinMax(d2Hw_mean_HobGho, force =T) #omg it doesnt auto display the actual min and max
Uniques<-terra::minmax(d2Hw_mean_HobGho)
Uniques.max <-Uniques["max","lyr.1"]
Uniques.min <-Uniques["min","lyr.1"]
#try a ggplot
d2Hw_mean_HobGho_plot <- ggplot()+
   geom_spatraster(data = d2Hw_mean_HobGho) +
   scale_fill_gradient2(limits = c(Uniques.min, Uniques.max), low = "#440154FF",mid= "#21908CFF", high = "#FDE725FF", midpoint=mean(c(Uniques.max,Uniques.min)),
    na.value = NA, breaks = seq(Uniques.min, Uniques.max, length.out = 5), labels = scales::label_number(accuracy =1)) +
   labs(fill= expression(paste(delta^{2},"H ", "(\u2030)")))+
   geom_sf(data=extent3,color="grey30",fill="transparent", lwd=0.3)+ #countries
   geom_spatvector(data=grat, color=alpha("grey60", 0.5)) +
   coord_sf(xlim = c(-10470815, 4682185),ylim = c(-1314163, 8259837), expand=F)+
   scale_x_continuous(breaks = seq(-160, 70, by = 40)) +
   scale_y_continuous(breaks = seq(0, 80, by = 20))+
   theme_minimal()+
   theme(plot.margin = margin(0, 0, 0, 0, "cm"),
         legend.margin=margin(0,0,.2,0),
         plot.background = element_rect(fill=NA, color=NA),
         panel.border=element_rect(fill=NA, color="grey30"),
         legend.background = element_rect(fill = NA, color=NA),
         legend.key.height = unit(1, "cm"),
         axis.title.x = element_blank(), 
     axis.title.y = element_blank())
ggsave(plot=d2Hw_mean_HobGho_plot, filename="isoscapes/d2Hw_mean_HobGho_plot.png", device="png", height=4, width=8, units="in",bg="white")      
```
```{r}
d2Hw_sd_HobGho <- terra::setMinMax(d2Hw_sd_HobGho, force =T) #omg it doesnt auto display the actual min and max
Uniques<-terra::minmax(d2Hw_sd_HobGho)
Uniques.max <-Uniques["max","lyr.2"]
Uniques.min <-Uniques["min","lyr.2"]
#try a ggplot
d2Hw_sd_HobGho_plot <- ggplot()+
   geom_spatraster(data = d2Hw_sd_HobGho) +
   scale_fill_gradient2(limits = c(Uniques.min, Uniques.max), low = "#0D0887FF",mid= "#CC4678FF", high = "#F0F921FF", midpoint=mean(c(Uniques.max,Uniques.min)),
    na.value = NA, breaks = seq(Uniques.min, Uniques.max, length.out = 5), labels = scales::label_number(accuracy =1)) +
   labs(fill= expression(paste(delta^{2},"H ", "(\u2030)")))+
   geom_sf(data=extent3,color="grey30",fill="transparent", lwd=0.3)+ #countries
   geom_spatvector(data=grat, color=alpha("grey60", 0.5)) +
   coord_sf(xlim = c(-10470815, 4682185),ylim = c(-1314163, 8259837), expand=F)+
   scale_x_continuous(breaks = seq(-160, 70, by = 40)) +
   scale_y_continuous(breaks = seq(0, 80, by = 20))+
   theme_minimal()+
   theme(plot.margin = margin(0, 0, 0, 0, "cm"),
         legend.margin=margin(0,0,.2,0),
         plot.background = element_rect(fill=NA, color=NA),
         panel.border=element_rect(fill=NA, color="grey30"),
         legend.background = element_rect(fill = NA, color=NA),
         legend.key.height = unit(1, "cm"),
         axis.title.x = element_blank(), 
     axis.title.y = element_blank())
ggsave(plot=d2Hw_sd_HobGho_plot, filename="isoscapes/d2Hw_sd_HobGho_plot.png", device="png", height=4, width=8, units="in",bg="white")      
```
Sr_mean,Sr_sd)
```{r}
Sr_mean <- terra::setMinMax(Sr_mean, force =T) #omg it doesnt auto display the actual min and max
Uniques<-terra::minmax(Sr_mean)
Uniques.max <-Uniques["max","rf_plantsoilmammal1"]
Uniques.min <-Uniques["min","rf_plantsoilmammal1"]
#try a ggplot
Sr_mean_plot <- ggplot()+
   geom_spatraster(data = Sr_mean) +
   scale_fill_gradient2(limits = c(Uniques.min, Uniques.max), low = "#440154FF",mid= "#21908CFF", high = "#FDE725FF", midpoint=mean(c(Uniques.max,Uniques.min)),
    na.value = NA, breaks = seq(Uniques.min, Uniques.max, length.out = 5), labels = scales::label_number(accuracy =.0001)) +
   labs(fill= expression(paste(~''^{87},"Sr/",~''^{86},"Sr")))+
   geom_sf(data=extent3,color="grey30",fill="transparent", lwd=0.3)+ #countries
   geom_spatvector(data=grat, color=alpha("grey60", 0.5)) +
   coord_sf(xlim = c(-10470815, 4682185),ylim = c(-1314163, 8259837), expand=F)+
   scale_x_continuous(breaks = seq(-160, 70, by = 40)) +
   scale_y_continuous(breaks = seq(0, 80, by = 20))+
   theme_minimal()+
   theme(plot.margin = margin(0, 0, 0, 0, "cm"),
         legend.margin=margin(0,0,.2,0),
         plot.background = element_rect(fill=NA, color=NA),
         panel.border=element_rect(fill=NA, color="grey30"),
         legend.background = element_rect(fill = NA, color=NA),
         legend.key.height = unit(1, "cm"),
         axis.title.x = element_blank(), 
     axis.title.y = element_blank())
ggsave(plot=Sr_mean_plot, filename="isoscapes/Sr_mean_plot.png", device="png", height=4, width=8, units="in",bg="white")      
```
```{r}
Sr_sd <- terra::setMinMax(Sr_sd, force =T) #omg it doesnt auto display the actual min and max
Uniques<-terra::minmax(Sr_sd)
Uniques.max <-Uniques["max","srse"]
Uniques.min <-Uniques["min","srse"]
#try a ggplot
Sr_sd_plot <- ggplot()+
   geom_spatraster(data = Sr_sd) +
   scale_fill_gradient2(limits = c(Uniques.min, Uniques.max), low = "#0D0887FF",mid= "#CC4678FF", high = "#F0F921FF", midpoint=mean(c(Uniques.max,Uniques.min)),
    na.value = NA, breaks = seq(Uniques.min, Uniques.max, length.out = 5), labels = scales::label_number(accuracy =.0001)) +
   labs(fill= expression(paste(~''^{87},"Sr/",~''^{86},"Sr")))+
   geom_sf(data=extent3,color="grey30",fill="transparent", lwd=0.3)+ #countries
   geom_spatvector(data=grat, color=alpha("grey60", 0.5)) +
   coord_sf(xlim = c(-10470815, 4682185),ylim = c(-1314163, 8259837), expand=F)+
   scale_x_continuous(breaks = seq(-160, 70, by = 40)) +
   scale_y_continuous(breaks = seq(0, 80, by = 20))+
   theme_minimal()+
   theme(plot.margin = margin(0, 0, 0, 0, "cm"),
         legend.margin=margin(0,0,.2,0),
         plot.background = element_rect(fill=NA, color=NA),
         panel.border=element_rect(fill=NA, color="grey30"),
         legend.background = element_rect(fill = NA, color=NA),
         legend.key.height = unit(1, "cm"),
         axis.title.x = element_blank(), 
     axis.title.y = element_blank())
ggsave(plot=Sr_sd_plot, filename="isoscapes/Sr_sd_plot.png", device="png", height=4, width=8, units="in",bg="white")      
```

# fin
```{r}
sessionInfo()
```

