---
title: "Suchan et al (in prep) - Isotope Rscript"
author: "Megan Reich"
date: "08/02/2022"
output: 
  html_document:
    theme: paper
    toc: true
    toc_float: true
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
```
# Purpose

Code for the isotope analyses section of:  
Suchan, T., Bataille, C.P., Reich, M.S., Toro-Delgado, E. Vila, R., Pierce, N.E., Talavera, G. (in prep.). A trans-oceanic aerial pathway for insect dispersal: combined evidence shows an 8,000 km journey for the butterfly Vanessa cardui.  
  
  
Isotope-based geographic assignment using hydrogen and strontium isotopes of 3 painted ladies captured in French Guiana by Gerard Talavera in October 2013. The hydrogen isotope values were measured in Utah by IsoForensics in June 2017. The strontium isotope ratios were measured at UBC by Megan Reich in December 2021.

# Setup
```{r message=FALSE, warning=FALSE}
if(!require(googledrive)) install.packages("googledrive"); library(googledrive)
if (!require(tidyverse)) install.packages(tidyverse); library(tidyverse)
if (!require(raster)) install.packages(raster); library(raster)
if (!require(rasterVis)) install.packages(rasterVis); library(rasterVis)
if (!require(assignR)) install.packages(assignR); library(assignR)
if (!require(rnaturalearth)) install.packages(rnaturalearth); library(rnaturalearth)
if (!require(viridis)) install.packages(viridis); library(viridis)
if (!require(lattice)) install.packages(lattice); library(lattice)
if (!require(unikn)) install.packages(unikn); library(unikn)
if (!require(ggpubr)) install.packages(ggpubr); library(ggpubr)
```
```{r}
#make some folders in home directory
dir.create("output")
dir.create("output/H")
dir.create("output/H_Sr")
dir.create("output/Sr")
dir.create("isoscapes")
```
# Data
```{r}
###Guiana isotope data
iso.data <- read.csv("unknown/IsoData_Guiana.csv") # tissue isotope data
```
```{r message=FALSE, warning=FALSE}
## define study extent extent 
box<-c(-120,50,-10,80)
extent1 <- ne_countries()
extent2<-crop(extent1,extent(box))
plot(extent2)

###Transform extent to isoscape projection
WGS84 <- CRS("+init=epsg:4326") 
eckert<-CRS("+proj=eck4 +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 +units=m +no_defs")
extent2<-spTransform(extent2,"+proj=eck4 +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 +units=m +no_defs")

###Create a world country shapefile
box1<-c(-180,180,-10,100)
extent3<-crop(extent1,extent(box1))
extent3<-spTransform(extent3,"+proj=eck4 +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 +units=m +no_defs")
```
```{r  warning=FALSE}
# Download Bataille et al 2020 Global bioavailable 87Sr/86Sr isoscapes from Google Drive:
drive_deauth()
drive_user()
mean_file <-  drive_get(as_id("11bMndaTM686gMycsZBnmdNCmJhI8M9l5"))
drive_download(mean_file, path= "isoscapes/Sr_Bataille2020_mean.tif")
sd_file <-  drive_get(as_id("1MfEU8-TQkvUokCqFFNv355_cuZUB8mpw"))
drive_download(sd_file, path= "isoscapes/Sr_Bataille2020_sd.tif")
```

```{r  warning=FALSE}
rf_sr<-raster("isoscapes/Sr_Bataille2020_mean.tif")
rf_sr_err <-raster("isoscapes/Sr_Bataille2020_sd.tif")

###Crop Sr isoscape to study area
Sr_europe <- crop(rf_sr, extent(extent2))
Sr_sd_europe <- crop(rf_sr_err, extent(extent2))
```
```{r}
print(iso.data[c(1:3,6,7,25,26)])
```
# d2H assignment

```{r  warning=FALSE}
###get Global mean-annual precipitation H and O isotope values
d2h_world<-assignR::getIsoscapes(isoType = "GlobalPrecipMA", timeout = 1200)
```
```{r  warning=FALSE}
###Reproject H isoscape to Sr isoscape
d2h_world2<-d2h_world #in case we want to do oxygen
d2h_world<-raster::subset(d2h_world,1:2)
d2h_world<-projectRaster(d2h_world,rf_sr,method="bilinear")
```
```{r  warning=FALSE}
###Cut wing H isoscape to same extent as Sr isoscape
d2h_world<-crop(d2h_world,Sr_europe)
d2h_world<-mask(d2h_world, Sr_europe)
```

The unknown origin data was measured in Utah (IsoForensics Inc) and is on Magozzi et al (2021) reference code: UT_H_2
```{r}
#Hobson 2019
caldata<-subOrigData(marker="d2H", taxon="Danaus plexippus", ref_scale="VSMOW_H") #ref_scale:	character string. Text identifier for reference scale to which all isotope values will be transformed. 
caldata
#Result: 150 samples from 31 sites Hobseon et al 2019
```
Make isoscape:
```{r }
d2H_wing<-calRaster(known=caldata, isoscape=d2h_world, interpMethod = 2, outDir="isoscapes")
```
```{r}
#add ref scale
iso.data$d2H_cal<-"UT_H_2"
#dataframe as required
samples<-iso.data %>%
  dplyr::select(c(d2H = d2h,d2H.sd = d2h_err,d2H_cal))
#transform to VSMOW
sample_trans<-refTrans(samples=samples, marker="d2H", ref_scale = "VSMOW_H")
```
```{r}
print(cbind(iso.data[,1:3], sample_trans$data))
```
```{r }
pdRaster(d2H_wing, unknown = sample_trans, genplot = T, outDir = "output/H/")
```
```{r }
#joint
joint_H<-assignR::jointP(H_postps)
writeRaster(joint_H, "output/H/Joint_H.tif")
```
```{r  warning=FALSE}
#reloading because pdRaster was run on a cluster
H_973<-raster("output/H/1_like.tif")
H_974<-raster("output/H/2_like.tif")
H_975<-raster("output/H/3_like.tif")
joint_H<-raster("output/H/Joint_H.tif")
```
```{r }
#normalize
Avg_H<-(H_973+H_974+H_975)/3
Avg_Hn<-Avg_H/cellStats(Avg_H,max) 
H_973n<-H_973/cellStats(H_973,max) 
H_974n<-H_974/cellStats(H_974,max) 
H_975n<-H_975/cellStats(H_975,max) 
joint_Hn<-joint_H/cellStats(joint_H,max) 
```
```{r}
#palette
pal2<-unikn::usecol(c("#F2F2F2","#21908CFF","#FDE725FF"), n=10) #10 colors
theme_vir<-rasterVis::rasterTheme(region=pal2)
theme_vir<-modifyList(theme_vir, list(axis.line=list(col="transparent")))
```
## joint and average probability
Average probability
```{r }
Avg_Hn_plot<-levelplot(Avg_Hn, maxpixels = 2e6, margin=FALSE, xlab=NULL, ylab=NULL, sub="normalized average probability", par.settings=theme_vir,
                scales = list(draw=FALSE), 
                at=seq(0,1, by=0.1), 
                colorkey=list(title=NULL, space="bottom", at=as.numeric(factor(c(seq(0,1, by=0.1)))), labels=as.character(c(seq(0,1, by=0.1))))) +
  latticeExtra::layer(sp.polygons(extent3, col="grey30", lwd=0.5)) #countries
Avg_Hn_plot$par.settings$layout.heights[
  c( 'top.padding',
     'key.sub.padding',
     'axis.xlab.padding',
     'key.axis.padding',
     'main.key.padding') ] <- .4
Avg_Hn_plot$par.settings$layout.widths[
  c( 'bottom.padding','left.padding',
     'right.padding') ] <- 1.3

Avg_Hn_plot<-ggarrange(Avg_Hn_plot)+bgcolor("white")
ggsave(plot=Avg_Hn_plot, filename="output/H/H_avg_FINAL.pdf", height=5, width=7, units="in")
ggsave(plot=Avg_Hn_plot, filename="output/H/H_avg_FINAL.png", height=3, width=4, units="in",bg="white")
```
Joint probability (product of probabilities)
```{r }
joint_Hn_plot<-levelplot(joint_Hn, maxpixels = 2e6, margin=FALSE, xlab=NULL, ylab=NULL, sub="normalized joint probability", par.settings=theme_vir,
                scales = list(draw=FALSE), 
                at=seq(0,1, by=0.1), 
                colorkey=list(title=NULL, space="bottom", at=as.numeric(factor(c(seq(0,1, by=0.1)))), labels=as.character(c(seq(0,1, by=0.1))))) +
  latticeExtra::layer(sp.polygons(extent3, col="grey30", lwd=0.5)) #countries
 joint_Hn_plot$par.settings$layout.heights[
  c( 'top.padding',
     'key.sub.padding',
     'axis.xlab.padding',
     'key.axis.padding',
     'main.key.padding') ] <- .4
joint_Hn_plot$par.settings$layout.widths[
  c( 'bottom.padding','left.padding',
     'right.padding') ] <- 1.3

joint_Hn_plot<-ggarrange(joint_Hn_plot)+bgcolor("white")
ggsave(plot=joint_Hn_plot, filename="output/H/H_joint_FINAL.pdf", height=5, width=7, units="in")
ggsave(plot=joint_Hn_plot, filename="output/H/H_joint_FINAL.png", height=3, width=4, units="in",bg="white")
```
## individual plots
```{r }
#try a lattice plot
H_973_plot<-levelplot(H_973n, maxpixels = 2e6, margin=FALSE, xlab=NULL, ylab=NULL, par.settings=theme_vir,
                scales = list(draw=FALSE),
                at=seq(0,1, by=0.1),
                colorkey=list(space="bottom", at=as.numeric(factor(c(seq(0,1, by=0.1)))), labels=as.character(c(seq(0,1, by=0.1))))) +
  latticeExtra::layer(sp.polygons(extent3, col="grey30", lwd=0.5)) #countries
H_973_plot$par.settings$layout.heights[
  c( 'bottom.padding',
     'top.padding',
     'key.sub.padding',
     'axis.xlab.padding',
     'key.axis.padding',
     'main.key.padding') ] <- .4
H_973_plot$par.settings$layout.widths[
  c( 'left.padding',
     'right.padding') ] <- 1.3


H_974_plot<-levelplot(H_974n, maxpixels = 2e6, margin=FALSE, xlab=NULL, ylab=NULL, par.settings=theme_vir,
                scales = list(draw=FALSE),
                at=seq(0,1, by=0.1),
                colorkey=list(space="bottom", at=as.numeric(factor(c(seq(0,1, by=0.1)))), labels=as.character(c(seq(0,1, by=0.1))))) +
  latticeExtra::layer(sp.polygons(extent3, col="grey30", lwd=0.5)) #countries
H_974_plot$par.settings$layout.heights[
  c( 'bottom.padding',
     'top.padding',
     'key.sub.padding',
     'axis.xlab.padding',
     'key.axis.padding',
     'main.key.padding') ] <- .4
H_974_plot$par.settings$layout.widths[
  c( 'left.padding',
     'right.padding') ] <- 1.3


H_975_plot<-levelplot(H_975n, maxpixels = 2e6, margin=FALSE, xlab=NULL, ylab=NULL, par.settings=theme_vir,
                scales = list(draw=FALSE), 
                at=seq(0,1, by=0.1),
                colorkey=list(space="bottom", at=as.numeric(factor(c(seq(0,1, by=0.1)))), labels=as.character(c(seq(0,1, by=0.1))))) +
  latticeExtra::layer(sp.polygons(extent3, col="grey30", lwd=0.5)) #countries
H_975_plot$par.settings$layout.heights[
  c( 'bottom.padding',
     'top.padding',
     'key.sub.padding',
     'axis.xlab.padding',
     'key.axis.padding',
     'main.key.padding') ] <- .4
H_975_plot$par.settings$layout.widths[
  c( 'left.padding',
     'right.padding') ] <- 1.3


joint_Hn_plot<-levelplot(joint_Hn, maxpixels = 2e6, margin=FALSE, xlab=NULL, ylab=NULL, par.settings=theme_vir,
                scales = list(draw=FALSE), 
                at=seq(0,1, by=0.1),
                colorkey=list(space="bottom", at=as.numeric(factor(c(seq(0,1, by=0.1)))), labels=as.character(c(seq(0,1, by=0.1))))) +
  latticeExtra::layer(sp.polygons(extent3, col="grey30", lwd=0.5)) #countries
joint_Hn_plot$par.settings$layout.heights[
  c( 'bottom.padding',
     'top.padding',
     'key.sub.padding',
     'axis.xlab.padding',
     'key.axis.padding',
     'main.key.padding') ] <- .4
joint_Hn_plot$par.settings$layout.widths[
  c( 'left.padding',
     'right.padding') ] <- 1.3

# save
H_plot<-ggarrange(H_973_plot, H_974_plot, H_975_plot, joint_Hn_plot, ncol=2, nrow=2)+bgcolor("white") #labels=c("A","B","C","D")
ggsave(plot=H_plot, filename="output/H/H_plot.pdf", height=10, width=14, units="in")
ggsave(plot=H_plot, filename="output/H/H_plot.png", height=6, width=8, units="in")
#end
```
# 87Sr/86Sr assignment
```{r }
###Cut wing Sr isoscape to same extent as H isoscape
Sr_europe<-crop(Sr_europe,d2h_world$d2h_MA) #hydrogen
Sr_europe<-mask(Sr_europe, d2h_world$d2h_MA)
Sr_sd_europe<-crop(Sr_sd_europe,d2h_world$d2h_MA) #hydrogen
Sr_sd_europe<-mask(Sr_sd_europe, d2h_world$d2h_MA)
###Create stack for assignR
Sr_stack<-stack(Sr_europe,Sr_sd_europe)

###dataset prepared for assignR
iso.data_Sr<-subset(iso.data, select=c("ID","Sr","Sr_err"))
```
```{r }
###pd Sr
assignR::pdRaster(Sr_stack, unknown = iso.data_Sr, outDir = "output/Sr/")
```
```{r }
###Joint Sr prob
Joint_Sr<-assignR::jointP(Sr_pd)
writeRaster(Joint_Sr, "output/Sr/Joint_Sr.tif")
```
```{r  warning=FALSE}
Sr_973<-raster("output/Sr/RVcoll14M973_like.tif")
Sr_974<-raster("output/Sr/RVcoll14M974_like.tif")
Sr_975<-raster("output/Sr/RVcoll14M975_like.tif")
Joint_Sr<-raster("output/Sr/Joint_Sr.tif")
```
```{r  warning=FALSE}
#normalize
Avg_Sr<-(Sr_973+Sr_974+Sr_975)/3
Avg_Srn<-Avg_Sr/cellStats(Avg_Sr,max) 
Sr_973n<-Sr_973/cellStats(Sr_973,max) 
Sr_974n<-Sr_974/cellStats(Sr_974,max) 
Sr_975n<-Sr_975/cellStats(Sr_975,max) 
Joint_Srn<-Joint_Sr/cellStats(Joint_Sr,max)
```
## joint and average probability
Average probability
```{r }
Avg_Srn_plot<-levelplot(Avg_Srn, maxpixels = 2e6, margin=FALSE, xlab=NULL, ylab=NULL, sub="normalized average probability", par.settings=theme_vir,
                scales = list(draw=FALSE), 
                at=seq(0,1, by=0.1), 
                colorkey=list(title=NULL, space="bottom", at=as.numeric(factor(c(seq(0,1, by=0.1)))), labels=as.character(c(seq(0,1, by=0.1))))) +
  latticeExtra::layer(sp.polygons(extent3, col="grey30", lwd=0.5)) #countries
Avg_Srn_plot$par.settings$layout.heights[
  c( 'top.padding',
     'key.sub.padding',
     'axis.xlab.padding',
     'key.axis.padding',
     'main.key.padding') ] <- .4
Avg_Srn_plot$par.settings$layout.widths[
  c( 'bottom.padding','left.padding',
     'right.padding') ] <- 1.3

Avg_Srn_plot<-ggarrange(Avg_Srn_plot)+bgcolor("white")
ggsave(plot=Avg_Srn_plot, filename="output/Sr/Sr_avg_FINAL.pdf", height=5, width=7, units="in")
ggsave(plot=Avg_Srn_plot, filename="output/Sr/Sr_avg_FINAL.png", height=3, width=4, units="in",bg="white")
```
Joint probability (product of probabilities)
```{r }
Joint_Srn_plot<-levelplot(Joint_Srn, maxpixels = 2e6, margin=FALSE, xlab=NULL, ylab=NULL, sub="normalized joint probability", par.settings=theme_vir,
                scales = list(draw=FALSE), 
                at=seq(0,1, by=0.1), 
                colorkey=list(title=NULL, space="bottom", at=as.numeric(factor(c(seq(0,1, by=0.1)))), labels=as.character(c(seq(0,1, by=0.1))))) +
  latticeExtra::layer(sp.polygons(extent3, col="grey30", lwd=0.5)) #countries
 Joint_Srn_plot$par.settings$layout.heights[
  c( 'top.padding',
     'key.sub.padding',
     'axis.xlab.padding',
     'key.axis.padding',
     'main.key.padding') ] <- .4
Joint_Srn_plot$par.settings$layout.widths[
  c( 'bottom.padding','left.padding',
     'right.padding') ] <- 1.3

Joint_Srn_plot<-ggarrange(Joint_Srn_plot)+bgcolor("white")
ggsave(plot=Joint_Srn_plot, filename="output/Sr/Sr_joint_FINAL.pdf", height=5, width=7, units="in")
ggsave(plot=Joint_Srn_plot, filename="output/Sr/Sr_joint_FINAL.png", height=3, width=4, units="in",bg="white")
```
## individual plots
```{r }
#try a lattice plot

Sr_973_plot<-levelplot(Sr_973n, maxpixels = 2e6, margin=FALSE, xlab=NULL, ylab=NULL, par.settings=theme_vir,
                scales = list(draw=FALSE), 
                at=seq(0,1, by=0.1),
                colorkey=list(space="bottom", at=as.numeric(factor(c(seq(0,1, by=0.1)))), labels=as.character(c(seq(0,1, by=0.1))))) +
  latticeExtra::layer(sp.polygons(extent3, col="grey30", lwd=0.5)) #countries
Sr_973_plot$par.settings$layout.heights[
  c( 'bottom.padding',
     'top.padding',
     'key.sub.padding',
     'axis.xlab.padding',
     'key.axis.padding',
     'main.key.padding') ] <- .4
Sr_973_plot$par.settings$layout.widths[
  c( 'left.padding',
     'right.padding') ] <- 1.3

Sr_974_plot<-levelplot(Sr_974n, maxpixels = 2e6, margin=FALSE, xlab=NULL, ylab=NULL, par.settings=theme_vir,
                scales = list(draw=FALSE),
                at=seq(0,1, by=0.1),
                colorkey=list(space="bottom", at=as.numeric(factor(c(seq(0,1, by=0.1)))), labels=as.character(c(seq(0,1, by=0.1))))) +
  latticeExtra::layer(sp.polygons(extent3, col="grey30", lwd=0.5)) #countries
Sr_974_plot$par.settings$layout.heights[
  c( 'bottom.padding',
     'top.padding',
     'key.sub.padding',
     'axis.xlab.padding',
     'key.axis.padding',
     'main.key.padding') ] <- .4
Sr_974_plot$par.settings$layout.widths[
  c( 'left.padding',
     'right.padding') ] <- 1.3

Sr_975_plot<-levelplot(Sr_975n, maxpixels = 2e6, margin=FALSE, xlab=NULL, ylab=NULL, par.settings=theme_vir,
                scales = list(draw=FALSE),
                at=seq(0,1, by=0.1),
                colorkey=list(space="bottom", at=as.numeric(factor(c(seq(0,1, by=0.1)))), labels=as.character(c(seq(0,1, by=0.1))))) +
  latticeExtra::layer(sp.polygons(extent3, col="grey30", lwd=0.5)) #countries
Sr_975_plot$par.settings$layout.heights[
  c( 'bottom.padding',
     'top.padding',
     'key.sub.padding',
     'axis.xlab.padding',
     'key.axis.padding',
     'main.key.padding') ] <- .4
Sr_975_plot$par.settings$layout.widths[
  c( 'left.padding',
     'right.padding') ] <- 1.3

Joint_Srn_plot<-levelplot(Joint_Srn, maxpixels = 2e6, margin=FALSE, xlab=NULL, ylab=NULL, par.settings=theme_vir,
                scales = list(draw=FALSE),
                at=seq(0,1, by=0.1),
                colorkey=list(space="bottom", at=as.numeric(factor(c(seq(0,1, by=0.1)))), labels=as.character(c(seq(0,1, by=0.1))))) +
  latticeExtra::layer(sp.polygons(extent3, col="grey30", lwd=0.5)) #countries
Joint_Srn_plot$par.settings$layout.heights[
  c( 'bottom.padding',
     'top.padding',
     'key.sub.padding',
     'axis.xlab.padding',
     'key.axis.padding',
     'main.key.padding') ] <- .4
Joint_Srn_plot$par.settings$layout.widths[
  c( 'left.padding',
     'right.padding') ] <- 1.3

#savey save
Sr_plot<-ggarrange(Sr_973_plot, Sr_974_plot, Sr_975_plot, Joint_Srn_plot, ncol=2, nrow=2)+bgcolor("white")
ggsave(plot=Sr_plot, filename="output/Sr/Sr_plot.pdf", height=10, width=14, units="in")
ggsave(plot=Sr_plot, filename="output/Sr/Sr_plot.png", height=6, width=8, units="in")
#end
```

# dual d2H - 87Sr/86Sr assignment

```{r }
###Prepare raster for dual assignments
dual_stack<-isoStack(d2H_wing,Sr_stack)
```
```{r}
###prepare dataset for dual assignment in assignR
iso.data<-cbind(iso.data, sample_trans$data)
iso.data_dual<-subset(iso.data, select=c("ID","d2H","Sr"))
iso.data_dual
```

```{r }
###Run dual assignments
dual_pd = assignR::pdRaster(dual_stack, unknown = iso.data_dual, outDir = "output/H_Sr/")
```
```{r }
#joint
joint_d<-assignR::jointP(dual_pd)
writeRaster(joint_d, "output/H_Sr/Joint_d.tif")
```

```{r  warning=FALSE}
d_973<-raster("output/H_Sr/RVcoll14M973_like.tif")
d_974<-raster("output/H_Sr/RVcoll14M974_like.tif")
d_975<-raster("output/H_Sr/RVcoll14M975_like.tif")
joint_d<-raster("output/H_Sr/Joint_d.tif")
```
```{r  warning=FALSE}
#normalize
Avg_d<-(d_973+d_974+d_975)/3
Avg_dn<-Avg_d/cellStats(Avg_d,max) 
d_973n<-d_973/cellStats(d_973,max) 
d_974n<-d_974/cellStats(d_974,max) 
d_975n<-d_975/cellStats(d_975,max) 
joint_dn<-joint_d/cellStats(joint_d, max)
```
## joint and average probability
Joint probability (product of probabilities)
```{r }
joint_dn_plot<-levelplot(joint_dn, maxpixels = 2e6, margin=FALSE, xlab=NULL, ylab=NULL, sub="normalized joint probability", par.settings=theme_vir,
                scales = list(draw=FALSE), 
                at=seq(0,1, by=0.1), 
                colorkey=list(title=NULL, space="bottom", at=as.numeric(factor(c(seq(0,1, by=0.1)))), labels=as.character(c(seq(0,1, by=0.1))))) +
  latticeExtra::layer(sp.polygons(extent3, col="grey30", lwd=0.5)) #countries
 joint_dn_plot$par.settings$layout.heights[
  c( 'top.padding',
     'key.sub.padding',
     'axis.xlab.padding',
     'key.axis.padding',
     'main.key.padding') ] <- .4
joint_dn_plot$par.settings$layout.widths[
  c( 'bottom.padding','left.padding',
     'right.padding') ] <- 1.3

joint_dn_plot<-ggarrange(joint_dn_plot)+bgcolor("white")
ggsave(plot=joint_dn_plot, filename="output/H_Sr/d_joint_FINAL.pdf", height=5, width=7, units="in")
ggsave(plot=joint_dn_plot, filename="output/H_Sr/d_joint_FINAL.png", height=3, width=4, units="in",bg="white")
```
Average probability
```{r }
Avg_d_plot<-levelplot(Avg_dn, maxpixels = 2e6, margin=FALSE, xlab=NULL, ylab=NULL, sub="normalized average probability", par.settings=theme_vir,
                scales = list(draw=FALSE), 
                at=seq(0,1, by=0.1), 
                colorkey=list(title=NULL, space="bottom", at=as.numeric(factor(c(seq(0,1, by=0.1)))), labels=as.character(c(seq(0,1, by=0.1))))) +
  latticeExtra::layer(sp.polygons(extent3, col="grey30", lwd=0.5)) #countries
 Avg_d_plot$par.settings$layout.heights[
  c( 'top.padding',
     'key.sub.padding',
     'axis.xlab.padding',
     'key.axis.padding',
     'main.key.padding') ] <- .4
Avg_d_plot$par.settings$layout.widths[
  c( 'bottom.padding','left.padding',
     'right.padding') ] <- 1.3

Avg_d_plot<-ggarrange(Avg_d_plot)+bgcolor("white")
ggsave(plot=Avg_d_plot, filename="output/H_Sr/d_avg_FINAL.pdf", height=5, width=7, units="in")
ggsave(plot=Avg_d_plot, filename="output/H_Sr/d_avg_FINAL.png", height=3, width=4, units="in",bg="white")
```
## individual plots
```{r }
#try a lattice plot
d_973_plot<-levelplot(d_973n, maxpixels = 2e6, margin=FALSE, xlab=NULL, ylab=NULL, par.settings=theme_vir,
                scales = list(draw=FALSE), 
                at=seq(0,1, by=0.1),
                colorkey=list(space="bottom", at=as.numeric(factor(c(seq(0,1, by=0.1)))), labels=as.character(c(seq(0,1, by=0.1))))) +
  latticeExtra::layer(sp.polygons(extent3, col="grey30", lwd=0.5)) #countries
d_973_plot$par.settings$layout.heights[
  c( 'bottom.padding',
     'top.padding',
     'key.sub.padding',
     'axis.xlab.padding',
     'key.axis.padding',
     'main.key.padding') ] <- .4
d_973_plot$par.settings$layout.widths[
  c( 'left.padding',
     'right.padding') ] <- 1.3

d_974_plot<-levelplot(d_974n, maxpixels = 2e6, margin=FALSE, xlab=NULL, ylab=NULL, par.settings=theme_vir,
                scales = list(draw=FALSE), 
                at=seq(0,1, by=0.1),
                colorkey=list(space="bottom", at=as.numeric(factor(c(seq(0,1, by=0.1)))), labels=as.character(c(seq(0,1, by=0.1))))) +
  latticeExtra::layer(sp.polygons(extent3, col="grey30", lwd=0.5)) #countries
d_974_plot$par.settings$layout.heights[
  c( 'bottom.padding',
     'top.padding',
     'key.sub.padding',
     'axis.xlab.padding',
     'key.axis.padding',
     'main.key.padding') ] <- .4
d_974_plot$par.settings$layout.widths[
  c( 'left.padding',
     'right.padding') ] <- 1.3

d_975_plot<-levelplot(d_975n, maxpixels = 2e6, margin=FALSE, xlab=NULL, ylab=NULL, par.settings=theme_vir,
                scales = list(draw=FALSE), 
                at=seq(0,1, by=0.1),
                colorkey=list(space="bottom", at=as.numeric(factor(c(seq(0,1, by=0.1)))), labels=as.character(c(seq(0,1, by=0.1))))) +
  latticeExtra::layer(sp.polygons(extent3, col="grey30", lwd=0.5)) #countries
d_975_plot$par.settings$layout.heights[
  c( 'bottom.padding',
     'top.padding',
     'key.sub.padding',
     'axis.xlab.padding',
     'key.axis.padding',
     'main.key.padding') ] <- .4
d_975_plot$par.settings$layout.widths[
  c( 'left.padding',
     'right.padding') ] <- 1.3

joint_dn_plot<-levelplot(joint_dn, maxpixels = 2e6, margin=FALSE, xlab=NULL, ylab=NULL, par.settings=theme_vir,
                scales = list(draw=FALSE), 
                at=seq(0,1, by=0.1),
                colorkey=list(space="bottom", at=as.numeric(factor(c(seq(0,1, by=0.1)))), labels=as.character(c(seq(0,1, by=0.1))))) +
  latticeExtra::layer(sp.polygons(extent3, col="grey30", lwd=0.5)) #countries
joint_dn_plot$par.settings$layout.heights[
  c( 'bottom.padding',
     'top.padding',
     'key.sub.padding',
     'axis.xlab.padding',
     'key.axis.padding',
     'main.key.padding') ] <- .4
joint_dn_plot$par.settings$layout.widths[
  c( 'left.padding',
     'right.padding') ] <- 1.3

# save
d_plot<-ggarrange(d_973_plot, d_974_plot, d_975_plot, joint_dn_plot, ncol=2, nrow=2)+bgcolor("white") #labels=c("A","B","C","D")
ggsave(plot=d_plot, filename="output/H_Sr/d_plot.pdf", height=10, width=14, units="in")
ggsave(plot=d_plot, filename="output/H_Sr/d_plot.png", height=6, width=8, units="in")
```
# Additional figures
## Isoscapes
```{r }
theme_vir<-rasterVis::rasterTheme(region=magma(10))
theme_vir<-modifyList(theme_vir, list(axis.line=list(col="transparent")))
theme_vir2<-rasterVis::rasterTheme(region=viridis(10))
theme_vir2<-modifyList(theme_vir2, list(axis.line=list(col="transparent")))

Hmean<-levelplot(d2H_wing$isoscape.rescale$mean, maxpixels = 2e6, margin=FALSE, xlab=NULL, ylab=NULL, par.settings=theme_vir2,
                scales = list(draw=FALSE) , 
                colorkey=list(space="bottom")) +
  latticeExtra::layer(sp.polygons(extent3, col="grey30", lwd=0.5)) #countries
Hmean$par.settings$layout.heights[
  c( 'bottom.padding',
     'top.padding',
     'key.sub.padding',
     'axis.xlab.padding',
     'key.axis.padding',
     'main.key.padding') ] <- .4
Hmean$par.settings$layout.widths[
  c( 'left.padding',
     'right.padding') ] <- 1.3


Hsd<-levelplot(d2H_wing$isoscape.rescale$sd, maxpixels = 2e6, margin=FALSE, xlab=NULL, ylab=NULL, par.settings=theme_vir,
                scales = list(draw=FALSE), 
                colorkey=list(space="bottom")) +
  latticeExtra::layer(sp.polygons(extent3, col="grey30", lwd=0.5)) #countries
Hsd$par.settings$layout.heights[
  c( 'bottom.padding',
     'top.padding',
     'key.sub.padding',
     'axis.xlab.padding',
     'key.axis.padding',
     'main.key.padding') ] <- .4
Hsd$par.settings$layout.widths[
  c( 'left.padding',
     'right.padding') ] <- 1.3


Smean<-levelplot(Sr_europe, maxpixels = 2e6, margin=FALSE, xlab=NULL, ylab=NULL, par.settings=theme_vir2,
                scales = list(draw=FALSE), 
                colorkey=list(space="bottom")) +
  latticeExtra::layer(sp.polygons(extent3, col="grey30", lwd=0.5)) #countries
Smean$par.settings$layout.heights[
  c( 'bottom.padding',
     'top.padding',
     'key.sub.padding',
     'axis.xlab.padding',
     'key.axis.padding',
     'main.key.padding') ] <- .4
Smean$par.settings$layout.widths[
  c( 'left.padding',
     'right.padding') ] <- 1.3


Ssd<-levelplot(Sr_sd_europe, maxpixels = 2e6, margin=FALSE, xlab=NULL, ylab=NULL, par.settings=theme_vir,
                scales = list(draw=FALSE), 
                colorkey=list(space="bottom")) +
  latticeExtra::layer(sp.polygons(extent3, col="grey30", lwd=0.5)) #countries
Ssd$par.settings$layout.heights[
  c( 'bottom.padding',
     'top.padding',
     'key.sub.padding',
     'axis.xlab.padding',
     'key.axis.padding',
     'main.key.padding') ] <- .4
Ssd$par.settings$layout.widths[
  c( 'left.padding',
     'right.padding') ] <- 1.3

iso_plot<-ggarrange(Hmean, Hsd, Smean, Ssd, ncol=2, nrow=2)+bgcolor("white") 
ggsave(plot=iso_plot, filename="isoscapes/isoscapes.pdf", height=10, width=14, units="in")
ggsave(plot=iso_plot, filename="isoscapes/isoscapes.png", height=6, width=8, units="in")
```
# fin
```{r}
sessionInfo()
```